### 2020-02-03 paper/pdot_table.tex
tweaked tables

\startlongtable
\begin{deluxetable*}{lllllllll}
%
\tablecaption{
	Predicted hot Jupiter period changes from linear radial
	velocity trends reported by \citet{knutson_friends_2014}.
\label{tab:pdot_table}
}
%
\tablenum{5}
%
% planet;gammadot;gammadot_pluserr;gammadot_minuserr;comment;pl_name;pl_orbper;Pdot;Pdot_upper_limit;Pdot_lower_limit;Pdot_perr;Pdot_merr;abs_Pdot;K14_significant
\tablehead{
  \colhead{Planet} &
  \colhead{$\dot{\gamma}$ [m$\,$s$^{-1}$$\,$yr$^{-1}$]} &
  \colhead{$+\sigma_{\dot{\gamma}}$ [m$\,$s$^{-1}$yr$^{-1}$]} & 
  \colhead{$-\sigma_{\dot{\gamma}}$ [m$\,$s$^{-1}$yr$^{-1}$]} & 
  \colhead{$P$ [days]} &
  \colhead{$\dot{P}_{\,{\rm RV}}$ [ms$\,$yr$^{-1}$]} &
  \colhead{$+\sigma_{\dot{P}_{\,{\rm RV}}}$ [ms$\,$yr$^{-1}$]} &
  \colhead{$-\sigma_{\dot{P}_{\,{\rm RV}}}$ [ms$\,$yr$^{-1}$]} &
  \colhead{Significant?}
}
%
%FIXME precision
\startdata
HAT-P-2 b & -0.0938 & 0.0067 & 0.0069 & 5.6335158 & -55.62 & 3.97 & 4.09 & 1
\enddata
%
\tablecomments{
  Table~5 is published in its entirety in a machine-readable format.
  The first entry is shown for guidance regarding form and content.
  Orbital periods were retrieved from NASA's Exoplanet Archive.
  Additional comments regarding non-linear trends and stellar activity
  are included in the MRT.
}
%
\end{deluxetable*}