### 2020-02-04 paper/ms.tex
Update ms.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt,twocolumn,tighten]{aastex62}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}

\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{\today}
\revised{---}
\accepted{---}
\submitjournal{AAS journals.}
\shorttitle{WASP-4 is accelerating towards the Earth}

\begin{document}

\defcitealias{bouma_wasp4b_2019}{B19}

\title{WASP-4 is Accelerating Towards the Sun}
% ", Explaining its Decreasing Orbital Period"
% ": A Cautionary Tale in the Search for Tidal Orbital Decay"
% "An Explanation for WASP-4's Decreasing Orbital Period".
% "WASP-4's Decreasing Orbital Period: Solved."

\correspondingauthor{L. G. Bouma}
\email{luke@astro.princeton.edu}

%
% key authors:
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}

%
% contributing authors: alphabetical
%
\author[0000-0001-8638-0320]{A. W. Howard}
\affiliation{Cahill Center for Astrophysics, California Institute of
Technology, Pasadena, CA 91125, USA}
%
\author{S. B. Howell}
\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
%
\author[0000-0002-0531-1073]{H. Isaacson}
\affiliation{Astronomy Department, University of California, Berkeley,
CA 94720, USA}
%
\author{H. Knutson}
\affiliation{Division of Geological and Planetary Sciences, California
Institute of Technology, Pasadena, CA 91125, USA}
%
\author{R. A. Matson}
\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
%

\begin{abstract}
  The orbital period of the hot Jupiter WASP-4b appears to be decreasing at
  a rate of $-8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$,
  based on transit-timing measurements spanning 12 years.
  Proposed explanations for the period change include
  tidal orbital decay, apsidal precession, and acceleration of the system
  along the line of sight.
  To investigate further, we performed new radial velocity measurements and
  speckle imaging of WASP-4.
  The radial-velocity data show that the system is
  accelerating towards the Sun at a rate of
  $-0.0422\pm 0.0028\,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.
  The associated Doppler effect would cause
  the apparent period to shrink at a rate of
  $-5.94 \pm 0.39~{\rm msec}\,{\rm yr}^{-1}$, comparable
  to the observed rate.
  Thus, the observed change in the transit period is mostly
  or entirely produced by the line-of-sight acceleration of the system.
  This acceleration is probably caused by a wide-orbiting companion
  of mass 10-200$\,M_{\rm Jup}$ and orbital distance 10-100$\,$AU,
  based on the magnitude of the radial-velocity trend and the non-detection of any
  companion in the speckle image.  We expect that 1 out of 6 hot Jupiters
  will show apparent period changes of the same order of magnitude as WASP-4,
  based on the hot-Jupiter companion statistics of  
  \citet{knutson_friends_2014}. Continued radial velocity
  monitoring of hot Jupiters is therefore essential to distinguish
  the effects of tidal orbital decay or apsidal precession from line-of-sight acceleration.
\end{abstract}

\keywords{Exoplanet tides (497), Exoplanet dynamics (490), Radial velocity 
(1332), Transit timing variation method (1710)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

The orbits of most hot Jupiters are formally unstable to tidal decay
\citep{counselman_outcomes_1973,
hut_stability_1980,rasio_tidal_1996,levrard_falling_2009,matsumura_tidal_2010}.
It is not clear, though, whether the timescale for tidal orbital decay is
shorter or longer than the timescale for main-sequence stellar
evolution.  This answer to this question depends on the uncertain rate at which
friction inside the star damps the tidal oscillations, which is ultimately
what causes the orbit to shrink
(as reviewed by \citealt{Mazeh2008} and
\citealt{ogilvie_tidal_2014}).
Population studies of hot Jupiters --- based on ages, rotation rates,
orbital distances, and Galactic kinematics --- have led to differing
conclusions, with estimated lifetimes ranging from less than a gigayear
to much longer than main-sequence lifetimes
(see, {\it e.g.},
\citealt{jackson_observational_2009}, \citealt{teitler_why_2014},
\citealt{penev_empirical_2018}, \citealt{cameron_hierarchical_2018},
\citealt{hamer_schlaufman_2019}).

An empirical resolution might be possible through long-term timing
of transits and occultations, seeking evidence for changes in the orbital
period.  For instance, long-term transit timing and
radial velocity measurements for WASP-12b have revealed a secular
decrease in the period at a rate of $\approx$30 milliseconds per year, which has
been interpreted as the effect of tidal orbital decay
\citep{maciejewski_departure_2016,patra_2017,yee_orbit_2020}.

This study draws attention to a confounding factor that, while elementary,
does not seem to have
received the attention it deserves.
The point is that observational programs
aimed at identifying orbital decay in hot Jupiters through transit
timing must be accompanied by concurrent long-term radial velocity
monitoring.  The reason is that an apparent change in period
can be produced by the Doppler effect associated with acceleration
of the hot-Jupiter host star along the line of sight, such as the acceleration that
might be produced by a massive, wide-orbiting companion.
Massive outer companions to hot Jupiters are common.
\citet{bryan_statistics_2016} calculated an
occurrence rate of $60.9^{+5.2}_{-5.6}\%$ for outer companions to hot
Jupiters with masses from 1-20$\,$$M_{\rm Jup}$ and semi-major axes
from 5-100$\,$AU.  Therefore, we expect that many hot Jupiters will display
secular trends in orbital period that are unrelated to tidal orbital decay.
This possibility can be checked by performing long-term radial-velocity monitoring
at a level sensitive enough to detect or rule out the relevant amplitude of
acceleration.

The focus of this study is the hot Jupiter WASP-4b, which has an
orbital period that appears to be decreasing by about 10 milliseconds
per year.  The period decrease was identified by \citet{bouma_wasp4b_2019}]{bouma_wasp4b_2019},
who combined data from
the NASA TESS mission \citep{ricker_transiting_2015} and a decade of ground-based
transit observations. Soon thereafter, \citet{southworth_transit_2019} reported an additional
22 transit times and recalculated the period derivative to be $\dot{P} = -9.2 \pm 1.1$
milliseconds per year.  A separate study by \citet{baluev_2019}
reported on additional transit times, and pointed out
that the period decrease was statistically significant only when
analyzing the data with the highest precision.

To determine the origin of the period change, we acquired four
additional radial velocity measurements with the Keck I 10m telescope
and the High Resolution Echelle
Spectrometer (HIRES; \citealt{vogt_hires_1994}).
In doing so, we extended
the time baseline of HIRES measurements from 3 to 9 years.
The previously available HIRES data led to the marginal ($\approx$$2\sigma$)
detection of a radial-velocity trend \citep{knutson_friends_2014}.
Our new measurements reveal a
line-of-sight acceleration of $\dot{\gamma} =
-0.0422^{+0.0028}_{-0.0027}\ {\rm m\,s^{-1}\,day^{-1}}$.
Through the Doppler effect\footnote{While the apparent period change caused by a
line-of-sight acceleration has been referred to as the ``R{\o}mer
effect'' \citep{yee_orbit_2020}, a simpler and probably better term is
the Doppler effect. The R{\o}mer effect refers to the {\it delay} in
the reception of a signal due to a change in the time required for
light to traverse the distance between the source and observer. The Doppler
effect refers to the change in the apparent {\it rate} of a process due
to changes in the relative motion of the source and observer, such
as the rate of transits.}, this translates into an expected period decrease
$-5.9$ milliseconds per year, which is comparable to the period decrease
that was measured from transit timing.  We undertook high-resolution
(speckle) imaging to search for evidence of a companion that could
be responsible for the acceleration of WASP-4.

Section~\ref{sec:observations} of this paper presents all of the
available transit data as well as the new radial velocity and
speckle imaging observations.  Section~\ref{sec:analysis} describes
our analysis of the data, and our interpretation
that WASP-4 is being pulled around by a brown dwarf or low-mass star.
Section~\ref{sec:discussion} places this
result within the context of orbital decay searches, and points
out that line-of-sight accelerations will be a relatively
common type of ``false positive.''
Section~\ref{sec:conclusions} offers concluding remarks.

\section{Observations}
\label{sec:observations}

\subsection{Transits}

Table~1 lists the transit times we collected for our analysis.  We
included data from the peer-reviewed literature for which {\it (i)} the
reported time was based on the data from a single transit (as opposed
to fitting the data spanning multiple transits and assuming the
period to be constant), {\it (ii)} the
central transit time was allowed to be a free parameter, and {\it (iii)} the time
system was documented clearly, in particular specifying
whether barycentric or heliocentric corrections had been
performed and whether leap seconds had been taken into account (TDB or UTC).

% FIXME : verify with Michael Gillon that this is true. (P. Evans
% done)
Most of these data are identical to the data presented by
\citetalias{bouma_wasp4b_2019}.  Twenty-two new times reported by
\citet{southworth_transit_2019} are included.  These transits were
observed from the 3.58m New Technology Telescope and the Danish 1.54m telescope at La Silla, Chile,
and the South African Astronomical Observatory 1.0m telescope.
% Five of the twenty-two occurred at simultaneous epochs as transits
% previously reported by \citet{hoyer_tramos_2013}.
Additional timing measurements were also reported recently by
\citet{baluev_2019}, based on a homogeneous analysis of archival
ground-based observations.
We included twelve of their ``high quality'' transit times from
the TRAPPIST telescope (six transits), El~Sauce (four transits), and
\citet{petrucci_no_2013}.  For TRAPPIST and El~Sauce, we verified with
the original observers that correct barycentric and leap-second
corrections had been performed (M.~Gillon, P.~Evans, priv{.} comm{.}).
We omitted the fourteen remaining \citeauthor{baluev_2019}
ETD\footnote{\url{http://var2.astro.cz/ETD}} times due to ambiguity in
whether leap-second corrections had been performed.
We did not include in our analysis the four occultation times tabulated by
\citetalias{bouma_wasp4b_2019}, because of the large timing
uncertainties and negligible statistical power.

\subsection{Radial velocities}

We acquired four new radial
velocity measurements with Keck/HIRES.  Our observations
were performed using the standard setup and reduction techniques of the
California Planet Survey \citep{howard_cps_2010}.  Previously, the
HIRES data-points spanned 2010 to 2013 \citep{knutson_friends_2014}.
Our new measurements triple the HIRES observing baseline to nine
years.

The complete set of radial velocity observations is given in Table~2.
Along with the 2010--2019 HIRES observations are early
measurements with two different spectrographs.
We included data from CORALIE ({\bf Spell it out on first use)}
from \citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010},
adopting the values from the homogeneous analysis of the latter authors.
We included data from HARPS ({\bf spell it out})
from \citet{pont_determining_2011} and \citet{husnoo_observational_2012}.
While \citet{triaud_spin-orbit_2010} also acquired HARPS data over
three nights for Rossiter-McLaughlin observations, these data were
reduced with a non-standard pipeline making them ill-suited
for our study, and we did not include them.\footnote{This problem was
fixed in principle by \citet{trifonov_public_2020} who performed
a homogeneous re-reduction of the entire HARPS data archive.
We found that the decision
regarding whether to include or omit these points did not
noticeably affect our results.}

\subsection{Speckle imaging}

Once we saw that the new HIRES observations implied a highly significant
trend in the radial velocity,
we sought independent evidence for a wide-orbiting companion by performing
speckle imaging with the Zorro instrument on the 
Gemini South 8m telescope\citep[see][and the instrument
web-pages\footnote{\url{www.gemini.edu/sciops/instruments/alopeke-zorro/}}]{scott_nessi_2018}.
Zorro is a dual-channel speckle interferometer employing narrow-band
filters centered at 562$\,$nm and 832$\,$nm.

We observed WASP-4 twice, on the night of September 11-12 with
relatively poor seeing (1.2$''$) and also on the night of September
28-29.  On each night, we acquired three sets of $1000\times 60$\,$msec exposures.
If a companion is present, the autocorrelation
functions of these speckle images would reveal a characteristic interference
pattern. This pattern in then used to determine the properties of the
detected companion and produce a reconstructed image.  Using the
reconstructed speckle images, contrast curves are produced to determine the
5-$\sigma$ detection limits (see \citealt{howell_speckle_2011}).  No
companions were detected. The data from the second night, which had better
seeing (0.6$''$), led to the most constraining limit.  The
832$\,$nm limits were the most useful, given that any faint
companion would likely be redder than the host star.  Therefore,
we opted to use
the 832$\,$nm September 28-29 contrast limits in the analysis described below.

\section{Analysis}
\label{sec:analysis}

\subsection{Transits}
\label{sec:transit_analysis}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.7\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
  The vertical axis shows the observed transit times minus the
  calculated times, assuming a constant orbital period.  More opaque
  points correspond to more precise data.  The $\pm1\sigma$
  uncertainties of the quadratic ephemeris 
  are shown in blue.
  The yellow star represents the weighted average of 18 data points
  obtained with TESS. The TESS data were averaged here for display purposes only;
  our analysis used the 18 individual transit times.
  \label{fig:times}
	}
\end{figure*}

We fitted two simple timing models to the transit-timing data. The first
model assumes the period $P$ to be constant:
\begin{align}
  t_{\rm tra}(E) &= t_0 + PE,
\end{align}
where $E$ is the integer specifying the number within the sequence of orbits
spanned by the data,
and $t_0$ is the transit time for the event designated $E=0$.
The second model assumes that the period changes at a steady
rate:
\begin{align}
  t_{\rm tra}(E) &=
    t_0 + PE +
    \frac{1}{2} \frac{{\rm d}P}{{\rm d}E} E^2.
  \label{eq:quadratic_model}
\end{align}
The free parameters are the reference time $t_0$, the period at the
reference epoch $P$, and the period derivative, ${\rm d}P/{\rm d}t =
(1/P) {\rm d}P/{\rm d}E$.  We defined the epoch numbers such that
$E=0$ is near the weighted average of the times of observed transits.
This choice leads to a small covariance between the uncertainties in $t_0$ and $P$.
%A third possible
%model that we did not consider for reasons that will become apparent
%is a precessing, eccentric orbit \citep[{\it
%e.g.},][]{gimenez_revision_1995,patra_2017}.

We fitted each model by assuming a Gaussian likelihood and
sampling over the posterior probability distributions.  We sampled the
posterior using the algorithm proposed by
\citet{goodman_ensemble_2010} and implemented by
\citet{foreman-mackey_emcee_2013} in \texttt{emcee}.  The prior for
the quadratic model allowed the period derivative to be either positive
or negative.

Figure~\ref{fig:times} shows the observed transit times, minus the
best-fitting constant-period model.  The best-fitting constant-period
model has 91 degrees of freedom, $\chi^2 = 276$, and $\chi^2_{\rm
red}=3.0$.  The best-fitting quadratic model has 90 degrees of freedom,
$\chi^2 = 183$, and $\chi^2_{\rm red}=2.0$.  The difference in the
Bayesian information criteria (BIC) between the linear and quadratic
and models is $\Delta {\rm BIC} = 89$, strongly favoring the
quadratic model \citep{kass_bayes_1995}.

From the high values of $\chi^2_{\rm red}$, we can surmise that neither model
provides a satisfactory fit to the transit data; there
must be some additional signal or noise.
The earlier study by \citetalias{bouma_wasp4b_2019} found that the quadratic model for the
(sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsening 
of $\chi^2_{\rm red}$ probably reflects underestimated measurement
uncertainties in at least some of the newly reported transit times,
due to time correlated noise in the photometry. It might also be due
to errors or misunderstandings of the time systems in which the
data were recorded, despite our best efforts to interpret the literature.
In what follows, we opted to enlarge the uncertainties in the best-fitting parameters
of each model by a factor of $(\chi^2_{\rm red})^{1/2}$.  This was a factor of $\approx$1.73$\times$ for the
linear model, and $\approx$1.41$\times$ for the quadratic model.

The resulting period derivative for the quadratic model is 
\begin{equation}
\dot{P}
  = - (2.74 \pm 0.28)\times 10^{-10}
  = - 8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}.
  \label{eq:dP_dt_obs}
\end{equation}
This agrees to within 1-$\sigma$ of the value reported by
\citet{southworth_transit_2019} ($\dot{P} = - 9.2 \pm 1.1~{\rm
ms}\,{\rm yr}^{-1}$).  It is 2.3-$\sigma$ larger than the rate
of period decrease reported by \citepalias[][; $- 12.6 \pm 1.2~{\rm ms}\,{\rm yr}^{-1}$]{bouma_wasp4b_2019},
presumably because of the new data from \citeauthor{southworth_transit_2019} and
\citeauthor{baluev_2019}. The other best-fitting model
parameters are reported in Table~3.

\subsection{Radial velocities}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.7\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Radial velocities of WASP-4.}
    {\it Top:} RV measurements, with best-fit instrument offsets
    added. 
    {\it Middle:}
    Residuals, after subtracting the
    maximum-likelihood orbit of WASP-4b.  The linear trend
    inferred from the RV data is shown with a black line.
    The trend that would be needed to
    produce the period decrease seen in transits ($\dot{P}= - 8.64 \pm
    1.26~{\rm msec}\,{\rm yr}^{-1}$) is indicated in purple.  The four
    new RV measurements from this work increase the
    significance of the linear trend from $\approx$2$\sigma$
    to 15$\sigma$.
    {\it Bottom:}
    Phased orbit of WASP-4b.
	\label{fig:rvs}
  \vspace{-0.0cm}
	}
\end{figure*}

Our initial model for the radial velocity data was a single Keplerian
orbit, plus instrument offsets, jitters, and a long-term linear trend
\citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
on the orbital period and time of inferior conjunction using the
values from Table~4 of \citetalias{bouma_wasp4b_2019}, and fixed
WASP-4b's eccentricity to zero
\citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
The free parameters were the velocity semi-amplitude, the instrument
zero-points, an additive ``white noise'' instrument jitter for each
instrument, and a linear ($\dot{v_{\rm r}}$) acceleration term.  We
also considered a model without the linear trend, and found that it
was disfavored by $\Delta {\rm BIC}=73$.

In the preferred ``planet+linear trend'' model, WASP-4 is accelerating
towards our line-of-sight at high confidence,
\begin{equation}
  \dot{v}_{\rm r} = \dot{\gamma} = 
     -0.0422^{+0.0028}_{-0.0027}
     \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}.
\end{equation}
The remaining radial velocity model parameters are listed in Table~4.
For comparison, before our new measurements, $\dot{\gamma}$ was
thought to be about five times smaller, and had marginal statistical
significance \citep{knutson_friends_2014,bouma_wasp4b_2019}.

The system's acceleration towards our line-of-sight causes a decrease
in the apparent orbital period:
\begin{align}
  \dot{P}_{\rm\,RV} &= \frac{\dot{v}_{\rm r} P}{c},
\end{align}
or in more convenient units,
\begin{align}
  \dot{P}_{\rm\,RV} &= 105{.}3\,{\rm msec}\,{\rm yr}^{-1} \
  \left( \frac{P}{{\rm day}} \right)
  \left( \frac{\dot{\gamma}}{ {\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}} \right).
\end{align}
For WASP-4, this yields
\begin{align}
  \dot{P}_{\rm\,RV} &= -5.94 \pm 0.39\,{\rm msec}\,{\rm yr}^{-1}.
\end{align}
The majority of the period decrease seen in transits ($\dot{P}= -
8.64 \pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) therefore seems to be caused by the
acceleration of the host star.

An important consideration is whether the measured RV trend is
correlated with stellar activity.  We investigate this by analyzing
WASP-4's emission in the Ca II H \& K lines, as quantified with the
chromospheric $S$-index \citep{wright_chromospheric_2004}.  We only
examined the HIRES velocities for this step, since they are the main
source of signal for our analysis.  First, we subtracted the orbital
solution from the Keck-HIRES velocities.  Then, following
\citet{bryan_statistics_2016,bryan_excess_2019}, we calculated the
Spearman rank correlation coefficient between the $S$-index and the
orbit-subtracted velocities.  We found a correlation coefficient of
0.16. This correlation is not statistically significant; the
corresponding $p$-value is 0.65.  Furthermore, inspection of the
$S$-index timeseries did not show secular or sinusoidal trends, as
would be expected if we were observing a long-term magnetic activity
cycle.  The $S$-index values are included in Table~2.  We conclude
that it is highly unlikely that the linear trend is caused by stellar
activity.

\subsection{Constraints on companion masses and semi-major axes}

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.47\textwidth]{f3.pdf}
	\end{center}
	\vspace{-0.7cm}
    \caption{
      {\bf Zorro contrast limits derived from point-source
      injection-recovery experiments.} Sources below the curve would
      have been detected.  The inset shows the speckle image
      reconstructed from 1000 60 millisecond frames in an 832$\,$nm
      bandpass, and acquired on September 28, 2019.  The image scale
      is $2.46''\times2.46''$.
    }
    \label{fig:zorro}
\end{figure}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.8cm}
	\caption{
    {\bf Masses and semi-major axes of companions that meet
    requirements of both the radial velocities and the speckle
    imaging.} The likelihood inferred from radial velocities is shown
    in grayscale, and the region excluded from the speckle imaging is
    indicated.  The expected mass to semimajor-axis degeneracy is
    shown with a black line.
	\label{fig:mass_sma}
  \vspace{-0cm}
	}
\end{figure*}

Given a linear radial velocity trend, we can place lower-limits on the
mass and semi-major axis of additional bodies in the system.  For a
quick estimate of the minimum mass required to explain the linear
trend in WASP-4, we turn to \citet{feng_california_2015}.  As they
discuss, the scenario that yields the minimum companion mass for a
system with a linear trend is a companion with $e\approx0.5$ and
$\omega=90^\circ$.  Substituting $P\approx 1.25\tau$ and $K \approx
0.5\tau \dot{\gamma}$ into the mass function \citep[{\it
e.g.},][]{wright_efficient_2009} yields
\begin{equation}
 M_{\rm min} \approx 5.99\,M_{\rm Jup}
  \left( \frac{\tau}{{\rm yr}} \right)^{4/3}
  \left| \frac{\dot{\gamma}}{{\rm m\,s^{-1}\,day^{-1}}} \right|
  \left( \frac{M_\star}{{M_\odot}} \right)^{2/3},
\end{equation}
where $\tau$ is the observing baseline.  For WASP-4, this gives
$M_{\rm min} = 4.9 M_{\rm Jup}$.  Higher masses are allowed for
companions that orbit further from the star: at fixed $\dot{\gamma}$,
$M_{\rm comp} \propto a^2$
\citep{torres_substellar_1999,liu_crossing_2002}.

High-resolution images can further limit the available parameter space
by setting an upper limit on the semimajor axis, and a maximum
brightness (and thereby mass) of any putative companions.  The
procedure we use to combine constraints from both radial velocities
and high resolution imaging has been developed by
\citet{wright_linear_trends_2007}, \citet{crepp_trends_2012},
\citet{montet_trends_2014}, \citet{knutson_friends_2014},
\citet{bryan_statistics_2016,bryan_excess_2019}, and others.

\paragraph{Speckle imaging constraints}

First, we would like to convert the contrast ratios obtained through
the Zorro imaging (Figure~\ref{fig:zorro}) to limits on the masses of
putative companions and their separations from the host star.

To do this, we followed \citet{montet_trends_2014}, and opted to
employ the \citet{baraffe_evolutionary_2003} models for substellar
mass objects and the MIST isochrones for stellar mass objects
\citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015,dotter_mesa_2016,choi_mesa_2016}.
We assumed that the system age was 5 Gyr, so that companions would
have fully contracted.

Due to the custom filters of the Zorro imager, and corresponding lack
of synthetic photometry, we further assumed that all sources had
blackbody spectra. While this is a simplification, we do not
readily have access to the planetary and stellar atmosphere models
needed for the consistent calculation with the \texttt{COND03} and
\texttt{MESA} models.  We therefore adopted the effective temperatures
and bolometric luminosities from the \citet{baraffe_evolutionary_2003}
and MIST isochrones.  Using these theoretical quantities and the
empirically-measured Zorro bandpasses, we calculated absolute
magnitudes in the 562 and 832 nm Zorro bands for stellar and planetary
mass companions.  Applying the same calculation to WASP-4 itself using
the effective temperature and bolometric luminosity from
\citetalias{bouma_wasp4b_2019}, we derived the transformation from
contrast ratio to companion mass.  The resulting limits are shown in
Figure~\ref{fig:mass_sma}.

\paragraph{Radial velocity constraints}

To derive constraints on possible companion masses and separations
from the radial velocities, we mostly followed the procedure of
\citet{bryan_excess_2019}.

We began by defining a $128\times128$ grid in true planetary mass and
semimajor axis, with even logarithmic spacing from 1 to 900$\,$$M_{\rm
Jup}$ and 3 to 500$\,$${\rm AU}$.  We then considered the possibility
that an additional companion in any particular cell could explain the
observed linear trend.  In each cell, we simulated 512 hypothetical
companions.

We assigned each companion a mass and semimajor axis from log-uniform
distributions within the grid cell. We drew the inclination from a
uniform distribution in $\cos i$.  For companion masses less than
$10\,M_{\rm Jup}$, we drew the eccentricity from
\citet{kipping_beta_2013}'s long-period exoplanet Beta distribution
($a=1.12$, $b=3.09$).  If the companion mass exceeded $10\,M_{\rm
Jup}$, we drew the eccentricity from the power-law $p_e \propto
e^\eta$ reported by \citet{moe_mind_2017} in their Equation~17 ($\eta
\approx 0.5$ for most orbital periods).  The long-period exoplanet and
long-period binary eccentricity distributions are quite different: the
exoplanet distribution is ``bottom-heavy'', with eccentricities
preferentially close to zero.  The binary star distribution is
``top-heavy'', with eccentricities closer to one.  We chose a mass
cutoff of $10\,M_{\rm Jup}$ to separate the two regimes based on the
bound observed by \citet{schlaufman_evidence_2018} between giant
planets and brown dwarfs, though this value is also close to the
$13\,M_{\rm Jup}$ deuterium-burning limit \citep[{\it
e.g.},][]{burrows_nongray_1997}.

For each simulated companion, we then drew a sample from the converged
chains of our initial model of WASP-4b. We subtracted the planet's
orbital solution, leaving RV points with a linear trend.  Given
$(a_{\rm c}, M_{\rm c}, e_{\rm c})$ for each simulated outer
companion, and the fixed instrument offsets and jitters from the MCMC
chains, we then performed a maximum likelihood fit for the time and
argument of periastron of the outer simulated companion.  We converted
the resulting $128\times128\times512$ cube of log-likelihood values to
probabilities, and averaged over the samples in each grid cell to
derive a probability distribution in mass and semi-major axis.
Figure~\ref{fig:mass_sma} shows the result.

\section{Discussion}
\label{sec:discussion}

\subsection{Implications for WASP-4}
Previous potential explanations for WASP-4b's decreasing orbital
period included tidal orbital decay, orbital precession, and
light-travel time effects \citep{bouma_wasp4b_2019}.  Our new radial
velocity measurements strongly indicate that the least exotic
option---light-travel time effects---is also the most likely.
Transits show the obital period decreasing by $- 8.64 \pm 1.26~{\rm
ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed in radial
velocities would predict a period decrease of $-5.94 \pm 0.39~{\rm
ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is $\approx$2$\sigma$
discrepant, Occam's razor would suggest that most or all of the apparent
decrease of WASP-4b's orbital period is caused by the line-of-sight
acceleration.

The corresponding requirements for the companion causing the
acceleration are that it is likely either a brown-dwarf or low mass
star, orbiting between $10$-$100\,{\rm AU}$ from the host star
(Figure~\ref{fig:mass_sma}).  Given such a mass, this companion could
at one time have influenced the orbital evolution of the inner giant.
The fact that most hot Jupiters have similar massive outer companions
\citep{knutson_friends_2014,bryan_statistics_2016} is circumstantial
evidence for certain high-eccentricity formation pathways (see
\citealt{dawson_johnson_2018}).  Further radial velocity monitoring
should eventually reveal the orbital parameters and minimum mass of
WASP-4's companion.

\subsection{How many other hot Jupiters are accelerating towards the
Earth?}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.4cm}
	\caption{
  {\bf Predicted hot Jupiter period changes from linear radial
  velocity trends.} Including WASP-4b, 16 of 51 hot Jupiters from
  \citet{knutson_friends_2014} have shown long-term radial velocity
  trends.  HAT-P-11 is shown, though its signal may be caused
  by stellar activity.  Three hot Jupiters are not shown because
  their radial velocity curves are better described as quadratic
  trends in time: HAT-P-17, WASP-8, and WASP-34.  Objects are ordered
  in the $y$ dimension by the absolute value of d$P$/d$t$.
	\label{fig:pdot_pop}
  \vspace{-0.3cm}
	}
\end{figure}

We identified WASP-4b's decreasing orbital period as part of a search
for tidal orbital decay.  However, most hot Jupiters have companions
outside of $5\,{\rm AU}$ with super-Jovian masses
\citep{knutson_friends_2014,bryan_statistics_2016}.  Line-of-sight
accelerations are correspondingly common in hot Jupiter systems.

To evaluate the importance of these effects for future transit timing
analyses, we collected the linear radial velocity trends reported by
\citet{knutson_friends_2014}, and computed the expected orbital period
derivatives $\dot{P}_{\rm\,RV} = \dot{v}_{\rm r} P/c$ for each system.
The results are given in Table~5, and visualized for hot Jupiters with
significant ($>$$3\sigma$) linear trends in Figure~\ref{fig:pdot_pop}.

Including WASP-4b, 16 of 51 hot Jupiters surveyed by
\citet{knutson_friends_2014} show a non-zero radial velocity trend.
Therefore around 1 in 3 hot Jupiters is expected to show period
changes commensurate with WASP-4 due to acceleration by outer
companions.  Half of these will be period decreases\footnote{ At a
population level, one would expect half of the line-of-sight
accelerations to be positive, and half negative.  Tidal orbital decay
however is usually negative, and could evenutally manifest as an asymmetry in the
$\dot{P}$ distribution of hot Jupiters.}, and will be an
astrophysical false positive in searches for tidal orbital decay.

\subsection{At what rate is the measurement precision of d$P$/d$t$
increasing?}
\label{sec:fisher}

For hot Jupiters that have been monitored over baselines
exceeding 10 years, secular changes in their orbital periods are
currently being constrained to a precision of $\lesssim$$10~{\rm
msec}\,{\rm yr}^{-1}$ ({\it e.g.}, K.~Patra et al{.} 2020, submitted).
This is roughly commensurate with the level of signal many outer
companions are expected to induce (Figure~\ref{fig:pdot_pop}).

It is therefore pertinent to ask at what point in time further
detections of the light-travel time effect will become routine for hot
Jupiters.  This question is the same as asking: at what rate does the
uncertainty in the quadratic term of Equation~\ref{eq:quadratic_model}
scale with the observing baseline?  We take a Fisher analysis approach
to the problem.  First, we rewrite Equation~\ref{eq:quadratic_model}
as
\begin{align}
  t_{\rm tra} = a_0 + a_1 E + a_2 E^2,
\end{align}
where $a_0\equiv t_0$, $a_1\equiv P$, and $a_2\equiv 0.5\cdot {\rm
d}P/{\rm d}E$.  Following \citet{gould_chi2_2003}, one can show that
if $N$ transit timing measurements are taken uniformly across a
baseline of $\Delta E$ epochs with constant precision $\sigma$, then
the uncertainty of the quadratic term is given by
\begin{align}
  \sigma_{a_2} = 6\sqrt{5}
   \frac{\sigma}{N^{1/2} (\Delta E)^2} \propto (\Delta E)^{-5/2}.
\end{align}
This result implies that a doubled observing baseline yields an
$\approx$5.7-fold improvement in precision on d$P$/d$t$.  If regular
observations continue from ground and space-based observatories,
period derivatives will be measured with precision exceeding
$1~{\rm msec}\,{\rm yr}^{-1}$ within the coming decade.

\section{Conclusions}
\label{sec:conclusions}

From newly acquired radial velocity measurements, we found that WASP-4
is accelerating towards the Earth at $\dot{\gamma} =
-0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}$.
The corresponding light-travel time effect predicts a period decrease
$\dot{\gamma} P/c$ of $-5.94 \pm 0.39~{\rm msec}\,{\rm yr}^{-1}$.  The
majority of the period decrease observed in transits ($\dot{P}= - 8.64
\pm 1.26~{\rm msec}\,{\rm yr}^{-1}$) is therefore explained by the
acceleration of the host star~---~not tidal orbital decay, or apsidal
precession.  The companion causing the acceleration is most likely a
brown dwarf or low-mass star with semi-major axis between
10-100$\,$AU.

Most hot Jupiters have outer companions with masses larger than
Jupiter beyond 5$\,$AU
\citep{knutson_friends_2014,bryan_statistics_2016}. The accelerations
and period changes induced by these outer companions will become an
increasingly large nuisance in the hunt for tidal orbital decay as the
observational baselines get longer.  In particular, the precision with
which the period derivative can be measured from transits scales with
the baseline duration to the power of 5/2 (Section~\ref{sec:fisher}), and
so within a decade many more hot Jupiters should show orbital period
changes due to accelerations from their outer companions.  
To distinguish this effect from tidal decay, further long-term radial
velocity measurements of hot Jupiters are strongly encouraged.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \acknowledgements
% %
% This paper includes data collected by the TESS mission, which are
% publicly available from the Mikulski Archive for Space Telescopes
% (MAST).
% %
% Funding for the TESS mission is provided by NASA's Science Mission
% directorate.
% %
% This work made use of NASA's Astrophysics Data System Bibliographic
% Services.
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% \newline
%

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  % \texttt{astroplan} \citep{astroplan2018},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  % \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
	\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{radvel} \citep{fulton_radvel_2018},
  % \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}

%
% The following are entries from Table 1 that are not otherwise cited
% in the text
%
\nocite{wilson_wasp-4b_2008}
\nocite{gillon_improved_2009}
\nocite{winn_transit_2009}
\nocite{hoyer_tramos_2013}
\nocite{dragomir_terms_2011}
\nocite{sanchis-ojeda_starspots_2011}
\nocite{nikolov_wasp-4b_2012}
\nocite{ranjan_atmospheric_2014}
\nocite{huitson_gemini_2017}

\input{WASP-4b_transit_time_table.tex}
\input{WASP-4b_rv_table.tex}
\input{model_fit_table.tex}
\input{rv_model_posterior_table.tex}
\input{pdot_table.tex}

%\clearpage
\bibliographystyle{yahapj}                            
\bibliography{bibliography}

\end{document}