### 2020-03-16 second_paper_review/response_20200315.txt
[WIP] second referee report

Dear Dr. Jackson,

We thank you for organizing the review of our paper, and thank the referee for
their helpful comments.  We have considered the referee's remarks carefully and
revised our manuscript accordingly.

Below, we have reproduced the referee's report, alongside with our
point-by-point responses. A list of changes is appended to the manuscript using
the trackchanges AASTeX macros.

Sincerely,

Luke Bouma

********************************************************************************
REFEREE COMMENT
----------
Section 2.1, Paragraph 2.

It would be helpful if the authors briefly describe the exact meaning of "high
quality" in the context of the transit times used in the analysis.

RESPONSE
----------
> Thank you for this prompt. We have removed the phrase "high quality", as its
> precise definition is not of particular relevance for this study. It is
> defined in Section 3.4 of Baluev+ 2019, through an empirical cutoff that
> Baluev+ made based on the sampling cadence and the light-curve residual RMS.

********************************************************************************
REFEREE COMMENT
----------
Section 3.1

In this section it is stated that the assumption of "a Gaussian likelihood and
sampling over the posterior probability distributions" is made. This assumption
needs further justification given the variety of data sources included. In
particular, it is later stated in Section 3.1 that the cause of the high
reduced chi2 values "must be some additional signal or noise". Alternatively,
it could indeed mean that the Gaussian likelihood assumption is not completely
valid.

RESPONSE
----------
> TODO

----------
********************************************************************************
REFEREE COMMENT
----------
Section 3.3

The discussion and figure in this section are very good. However, it is not
sufficiently demonstrated that a stellar companion is ruled out when
considering the possible angular separations of the companion resulting from
being near inferior or superior conjunction. This is particularly important for
this system since even a small degree of coplanarity of the companion with
WASP-4b would result in an edge on orbit that makes the companion difficult to
detect when near conjunction locations. Also, the very last sentence of this
section "Figure 4 shows the result." should describe the results of the figure
in more detail.

RESPONSE
----------
> TODO
>
> The comment regarding the last sentence has been addressed.

********************************************************************************
REFEREE COMMENT
----------
Section 4.1

Paragraph 1. The paragraph implies that the Doppler effect is the sole cause of
the timing derivative. However, and given the 2 sigma discrepancy, the wording
should allow for the possibility that the timing derivative is actually a
result of a combination of the described effects.

RESPONSE
----------
> We have modified the wording in this paragraph to clarify that this is a
> possibility, but that at present the data are best explained

********************************************************************************
REFEREE COMMENT
----------
Section 4.2

Paragraph 3. "hot Jupiters is expected" should be "hot Jupiters are expected".

RESPONSE
----------
> Corrected, thank you.